---
title: "DataTidying"
author: "Benjamin Keisling"
date: "1/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=FALSE, message=FALSE}

library(dplyr)
library(tidyr)
library(ggplot2)

```

# Data Cleaning

Read in Data File

```{r}
catch_df <- read.csv(file = "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1",
                     stringsAsFactors = FALSE)

head(catch_df)
```

cmd + shift + m %>% pipe operator


In the following functional chunk, we removed two of the columns (marginal sum and notes). The we gathered all the data by adding a species key, but kept the region and year. Now each column represents a single observation. We have converted the data from a "wide" format (lots of columns) to a long format (lots of rows).

```{r}

catch_long <- catch_df %>%
  select(-All,-notesRegCode) %>% 
  gather(key = "species", value = "catch",-Region,-Year) 
summary(catch_long)
head(catch_long)
```

Note that the summary shows there are some errors in the dataset, because it included the "catch" column in the "Character" class whereas it should be numeric. So I found the bad character, which was an "I" but it was supposed to be a one. So I replaced it. 

```{r}
catch_cleaned <- catch_long %>%
  rename(catch_thousands = catch) %>%
  mutate(catch_thousands = ifelse(catch_thousands == "I", "1", catch_thousands)) %>% 
  mutate(catch_thousands = as.integer(catch_thousands)) %>% 
  mutate(catch = catch_thousands * 1000)

head(catch_cleaned)
```


```{r, eval=FALSE, echo = FALSE}
test_catch <- as.integer(catch_cleaned$catch_thousands)

i <- which(is.na(test_catch)==TRUE)

j <- which(test_catch>=25000)

catch_cleaned[j,]
```

Now the data is ready for analysis.

# Split-Apply-Combine

```{r}
catch_total <- catch_cleaned %>%
  filter(catch>0) %>% 
  group_by(Year,Region) %>% 
#  summarize(catch_region = mean(catch))

  summarize(catch_region = mean(catch), n_obs = n(), stdev_catch = sd(catch))

catch_total
#head(catch_cleaned)
```


Filter for different kinds of salmon and rearrange the data.

```{r, eval=FALSE,echo=FALSE}
catch_chinook <- catch_total %>% 
  filter(species == "Chinook" | species == "Chum") %>% 
  arrange(-Year)
head(catch_chinook)
#plot(catch_total)
```



# Joins

Read in a second table to practice joins.
```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                        stringsAsFactors = FALSE)
head(region_defs)
```


```{r}
region_clean <- region_defs %>% 
  select("code","mgmtArea") %>% 
  rename("Region" = "code")
  #remove the columns we don't need
head(region_clean)
```


```{r}

catch_joined <- left_join(catch_cleaned,region_clean)
head(catch_joined)

```


Learning how to spread data.

```{r}
catch_wide <- catch_cleaned %>% 
  filter (Year > 1980) %>% 
  select(-catch_thousands) %>% 
  spread(key = Year, value = catch) #%>% 
 # select(-Region,-species)
head(catch_wide)
#heatmap(catch_wide)
```


ok now we're doing unite and separate

```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)
```


reformat the dates...

```{r}
dates_sep <- dates_df %>% 
  separate(col= date, int = c("month", "day", "year"), sep ="/", remove = 1)

head(dates_sep)
```

now we unite in the format we like...

```{r}
dates_unite <- dates_sep %>%
  unite(date_iso, "year", "month", "day", sep = "-", remove=TRUE)

head(dates_unite)
  
# @Ropensci, @Rstats can help with problems

```


for dates you can you the LUBRIDATE package.... the name of this... skjsjksjskjskjskjs
