---
title: "DataTidying"
author: "Benjamin Keisling"
date: "1/16/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, warning=FALSE, message=FALSE}

library(dplyr)
library(tidyr)
library(ggplot2)

```

# Data Cleaning

Read in Data File

```{r}
catch_df <- read.csv(file = "https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1",
                     stringsAsFactors = FALSE)

head(catch_df)
```

cmd + shift + m %>% pipe operator


In the following functional chunk, we removed two of the columns (marginal sum and notes). The we gathered all the data by adding a species key, but kept the region and year. Now each column represents a single observation.

```{r}

catch_long <- catch_df %>%
  select(-All,-notesRegCode) %>% 
  gather(key = "species", value = "catch",-Region,-Year) 
summary(catch_long)
head(catch_long)
```

Note that the summary shows there are some errors in the dataset, because it included the "catch" column in the "Character" class whereas it should be numeric.





```{r}
catch_cleaned <- catch_long %>%
  rename(catch_thousands = catch) #%>%
 # mutate(catch = catch_thousands * 1000)
  
head(catch_cleaned)
```


```{r}
test_catch <- as.integer(catch_cleaned$catch_thousands)

i <- which(is.na(test_catch)==TRUE)

j <- which(test_catch>=25000)

catch_cleaned[j,]
```



# Split-Apply-Combine

# Joins