---
title: "index"
author: "Benjamin Keisling"
date: "1/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load in libraries
```{r, warning = F, message = F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
```

Read in data from: Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.

```{r}
esc <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", stringsAsFactors = F)
head(esc)
```

#Data Tidying

Calculate annual, regional, total escapement

* `separate` the data into year - month - day
* `group_by` `summarize` to calculate the sum over unique region-year-species combinations
* `filter` for salmon species

```{r}
annual_esc <- esc %>% 
  separate(sampleDate, into = c("year","month","day"), sep = "-") %>% 
  mutate(year = as.numeric(year)) %>% 
  group_by(year,SASAP.Region,Species) %>%
  summarize(escapement = sum(DailyCount), n_obs = n()) %>% 
  filter(Species %in% c("Chinook","Sockeye","Pink","Chum","Coho"))


head(annual_esc)
```

Now we have the total count by year and region. Creating an interactive Data Table.

```{r}
datatable(annual_esc)
```




#Static Plots

Now were gonna make a plot. Note that in the AESTHETICS (aes) call, everything should refer to a column in your data.

RColorbrewer. tons of GGplot palettes.

timeseries of Kodiak escapement by species
```{r}

mytheme <- theme_light() + theme(plot.title = element_text(hjust=1)) + 
  theme(plot.subtitle = element_text(hjust=1))


ggplot(data=filter(annual_esc, SASAP.Region=="Kodiak"), aes(x = year, y=escapement, color=Species)) + geom_line() + geom_point() + ylab(expression(Escapement~(number~of~fish)~W/m^2)) + xlab("Year") + ggtitle("Kodiak Salmon Escapement","Organized by Species") + mytheme

```


```{r}

ggplot(annual_esc, aes(x = Species, y = escapement, fill = Species)) + geom_col()

```

```{r}

ggplot(annual_esc, aes(x = Species, y = escapement, fill=SASAP.Region)) + geom_col()

```




```{r}
ggplot(filter(annual_esc, SASAP.Region == "Kodiak", Species == "Chum" | Species == "Coho" | Species == "Chinook"), aes(x = year, y = escapement, color = Species)) + 
    geom_line() +
    geom_point() +
    ylab("Escapement") +
    ggtitle("Kodiak Salmon Escapement") +
    theme_bw()
```


```{r}
ggplot(filter(annual_esc, SASAP.Region == "Kodiak", Species == "Coho"), aes(x = year, y = escapement, color = escapement)) + 
    geom_line() +
    geom_point() +
    ylab("Escapement") +
    ggtitle("Kodiak Salmon Escapement") +
    theme_bw()
```

#Interactive Map
