---
title: "index"
author: "Benjamin Keisling"
date: "1/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

Load in libraries
```{r, warning = F, message = F, echo = F}
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(leaflet)
```

Read in data from: Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M.

```{r}
esc <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", stringsAsFactors = F)
head(esc)
```

#Data Tidying

Calculate annual, regional, total escapement

* `separate` the data into year - month - day
* `group_by` `summarize` to calculate the sum over unique region-year-species combinations
* `filter` for salmon species

```{r}
annual_esc <- esc %>% 
  separate(sampleDate, into = c("year","month","day"), sep = "-") %>% 
  mutate(year = as.numeric(year)) %>% 
  group_by(year,SASAP.Region,Species) %>%
  summarize(escapement = sum(DailyCount), n_obs = n()) %>% 
  filter(Species %in% c("Chinook","Sockeye","Pink","Chum","Coho"))


head(annual_esc)
```

Now we have the total count by year and region. Creating an interactive Data Table.

```{r}
datatable(annual_esc)
```




#Static Plots

Now were gonna make a plot. Note that in the AESTHETICS (aes) call, everything should refer to a column in your data.

RColorbrewer. tons of GGplot palettes.

timeseries of Kodiak escapement by species
```{r}

mytheme <- theme_classic() + theme(plot.title = element_text(hjust=1)) + 
  theme(plot.subtitle = element_text(hjust=1))


ggplot(data=filter(annual_esc, SASAP.Region=="Kodiak"), aes(x = year, y=escapement, color=Species)) + geom_line() + geom_point() + ylab(expression(Escapement~(number~of~fish)~W/m^2)) + xlab("Year") + ggtitle("Kodiak Salmon Escapement","Organized by Species") + mytheme

```


```{r}

ggplot(annual_esc, aes(x = Species, y = escapement, fill = Species)) + geom_col()

```

```{r}

ggplot(annual_esc, aes(x = Species, y = escapement, fill=SASAP.Region)) + geom_col()

```




```{r}
ggplot(filter(annual_esc, SASAP.Region == "Kodiak", Species == "Chum" | Species == "Coho" | Species == "Chinook"), aes(x = year, y = escapement, color = Species)) + 
    geom_line() +
    geom_point() +
    ylab("Escapement") +
    ggtitle("Kodiak Salmon Escapement") +
    theme_bw()
```


```{r}
ggplot(filter(annual_esc, SASAP.Region == "Kodiak", Species == "Coho"), aes(x = year, y = escapement, color = escapement)) + 
    geom_line() +
    geom_point() +
    ylab("Escapement") +
    ggtitle("Kodiak Salmon Escapement") +
    theme_bw()
```


Plot timeseries of escapement by species for all regions

```{r, fig.height=6, fig.width=10, echo = FALSE}

mytheme <- theme_classic() + theme(plot.title = element_text(hjust=1)) + 
  theme(plot.subtitle = element_text(hjust=1))

ggplot(annual_esc, aes(x = year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  facet_wrap(~SASAP.Region, scales = "free_y") +
  ylab("escapement (number of fish)") +
  mytheme

#ggsave("region_escapement.pdf") # PNG, JPEG (raster) or SVG/PDF/ESP (vector for editing after)
```

Now what if we want to save the data that we have cleaned up and modified?

```{r}
write.csv(annual_esc,"annual_escapement.txt", row.names = F)
```


#Interactive Map

Make a lap using leaflet

* find unique locations with lat/long values

```{r}
locations <- esc %>% 
  drop_na() %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na()

datatable(locations)
```

Now that we have the locations we can create a leaflet map.

```{r}
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(~ Longitude, ~ Latitude, popup= ~ Location)

#package called SF or geomSF allows you to make maps using GGplot. 
#you have to set leaflet to use the projection that is within the tiles you are using. based on your basemap. you can use ESRI's geospatial network. so you just have to look into these tile server things...
```



